Active Object Search with a Mobile Device for the Visually Impaired

### 2018-11-22 paper.tex
Removed confusing figure, changed 2 plots to clarify info, implemented changes discussed in office
sec I and II revised

\abstract{Smartphones can be used to provide a multitude of services to assist people with visual impairments. The cameras embedded in these mobile devices, in particular, can be used for many useful tasks, including searching particular objects in unknown environments. Although previous research in computer vision has looked at ways to solve these tasks by processing the video frames of the mobile camera, very little work has been done in {\em actively} guiding the user towards specific point of interest that maximise the effectiveness of the underlying visual algorithms.
In this paper, we propose a control algorithm based on a Markov Decision Process (MDP) that generates instructions, in real-time, to guide a user searching for a target object with a smartphone's camera. The solution is part of a more general active vision application for the visually impaired. An initial implementation of the system on an Android smartphone was experimentally evaluated with sighted participants to determine the performance of the control algorithm. The results show the effectiveness of our solution and its potential application to help the visually impaired find objects in unknown environments.}

@@

\noindent It is estimated that almost half a billion people worldwide live with mild to severe visual impairments or total blindness~\cite{bourne2017magnitude} and significant effort is being made to enable these people to lead more normal and independent lives. Modern improvements to mobile computing power and image processing techniques have provided researchers with new and powerful tools to solve this problem.
%The ActiVis project aims to contribute to this effort by delivering an indoor navigation system targeted towards enabling blind and visually impaired (VI) users to safely navigate towards their desired destination or
The work presented here is part of a project to assist the visually impaired (VI) navigate and find objects in unknown environments with the aid of a smartphone. The proposed system implements ideas from the field of active vision~\cite{Bajcsy2017}, but replaces the typical electro-mechanical actuators of the moving camera with the body (i.e.\ arm, hand) of the user holding the smartphone, as pictured in Figure~\ref{fig:system-use}. Schematically, the actuators used in a typical control loop, pictured in Figure~\ref{fig:control-loop}, are replaced by a human. %This project expands upon concepts originally proposed in~\cite{bellotto2013} and~\cite{lock2017portable}.

\begin{figure}
  \centering
  \begin{subfigure}
    \centering
    \includegraphics[width=0.38\textwidth]{figures/system_use.png}
    \caption{The system in use during an experiment. }\label{fig:system-use}
  \end{subfigure}
  \hspace{8mm}
  \begin{subfigure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/control_loop.png}
    \caption{The system layout and control loop. $r$ is the reference object, $e$ the error signal, $u$ and $u^*$ the original and interpreted control signals and $y$ is the current object observation. $K$, $H$ and $P$ are the control, human and sensor blocks respectively. }\label{fig:control-loop}
  \end{subfigure}
\end{figure}

The goal of this system is to understand the user's surroundings and determine what the next best course of action is to reach the target location based on what is currently within view and what has been observed in the past. We trained a Markov Decision Process (MDP)~\cite{bellman1957markovian} to produce a so-called policy that defines the optimal action the agent (the guidance system in this case) takes when it finds itself within some state.

This work's contributions are

\begin{itemize}
  \item a policy-driven human controller that can drive a user towards a point in space,
  \item a data-based MDP model that understands the spatial relationship between objects and
  \item experiments that prove the effectiveness of the system implementation.
\end{itemize}

@@

Section~\ref{sec:previous-work} discusses other relevant work done in this field, followed by a detailed discussion on the design of the guidance system, human-control module and MDP\@ in Sections~\ref{sec:system-design} and~\ref{sec:controller-design}. Then we discuss the experiment and present its results in Section~\ref{sec:experiments}, after which we conclude the paper and discuss the project's next steps in Section~\ref{sec:conclusion}.

@@

\noindent Assistive vision and guidance for the VI is an active and diverse field of study. In recent years, the increase in mobile processing power and computer vision improvements have led to researchers investigating using a cellphone camera to augment or enhance a user's vision and help them find objects or another point of interest. Earlier attempts at the problem involved placing special markers or barcodes, encoded with information about an object or the area, around an environment, which the user then scans with a cellphone or similar mobile device~\cite{gude2013blind,iannizzotto2005badge3d,manduchi2012mobile}. This device then uses some feedback mode, e.g.\ Braille and sound, to guide the user towards the target. %The major drawbacks to this approach is the significant upfront effort required to augment entire areas with markers to a point where the system will be useful to a VI user. Furthermore, these passive systems rely on the user scanning the marker in the first place; something that can be made more complicated with markers degrading over time, changing ambient lighting, different viewing angles and marker occlusion.

Another approach is to discard tags completely and rely on computer vision to perform the object detection, something that has become more plausible with recent improvements to feature detectors and deep networks~\cite{huang2017speed,redmon2016you}. SIFT and SURF-based object detectors have also been used to detect known objects when they are in the camera's view and guide the user to them using sonified guidance instructions~\cite{schauerte2012assistive}. This system is more flexible than the tag-based ones, but have the same drawback of being passive and rely on chance for the user to place the object within the camera's view in the first place. The paper also does not report on any performance metrics. The VizWize system~\cite{bigham2010vizwiz} offloads the object recognition tasks to an Amazon Mechanical Turk worker that then provides feedback on where the object of interest is located relative to the user. The VizWiz has the advantage of being fairly robust and is able to classify a great deal of objects with little effort from the user and can provide natural, human-generated and curated directions. However, this approach does not address the issue of enhancing user independence, since the user is now beholden to an online worker instead of a relative, friend or bystander. Furthermore, a good internet connection is required on the device, perhaps limiting use in some poor-reception areas.

Researchers implemented active search and perception~\cite{Bajcsy2017} strategies in robots and image classifiers in an attempt to optimise their classification and planning tasks by exploiting the structured nature of human object placements.

Two research teams have recently implemented an active object search strategy into their image classifiers~\cite{caicedo2015active,gonzalez2015active}. Their approaches use different methods, but both involve using a model to generate windows of interest to submit to a classifier for classification. The size and locations of the windows within the image are generated using the spatial relationship between objects, taken from the SUNCG and PASCAL datasets, and are iteratively changed based on the output from the respective models. The advantage of their approaches over their contemporaries is that far fewer windows are generated and submitted to the classifier, resulting in lower object classification times. They also report comparable accuracy compared to their counterparts.

Similarly, roboticists have moved to incorporate similar strategies on robotic platforms to improve autonomous object search, manipulation and localisation tasks. Researchers have developed a planning algorithm that allows a robotic manipulator to select an object to move that optimally executes an object search or exploration task in a cluttered environment~\cite{dogar2014object}. Another team implemented an MDP that generates the optimal search strategy to find an object in a room over a belief state of object positions and configurations~\cite{aydemir2011search}. However, the authors trained their MDP using a custom object-placement and configuration probabilities and have shown that the results are sensitive to changes within this distribution.

In all, much research has been conducted on object recognition and guidance towards an object once detected, as well as active vision within image classifiers and robotic systems. However, work on active object search and guidance with a human is harder to come by. In this work we attempt to implement an active vision guidance system with a human in the loop~\cite{bellotto2013,lock2017portable} that can guide a user towards an out-of-view target object by using prior knowledge of the spatial distribution of objects within an indoor environment and the history of past object observations made during the search. With this, we hope to make the searching process less random and move towards a more natural and efficient object searching process to the benefit of the VI\@.

@@

In comparison to the VizWiz system~\cite{bigham2010vizwiz} (mean 92s, standard deviation 37.7s), our results look very reasonable. However, these results may greatly change when we test the system with VI participants like they did in~\cite{bigham2010vizwiz}.

\todo{COMPARE RESULTS FROM ROBOTS? APPROX 100s FOR PLANNING AND EXECUTION}

>>>>

\abstract{Modern smartphones can provide a multitude of services to assist people with visual impairments. The cameras embedded in these mobile devices, in particular, can be used for many useful tasks, like reading signs or searching for objects in unknown environments. Although previous research in computer vision has looked at ways to solve these tasks by processing the video frames of the mobile camera, very little work has been done in {\em actively} guiding the user towards specific points of interest that maximise the effectiveness of the underlying visual algorithms.
In this paper, we propose a control algorithm based on a Markov Decision Process (MDP) that generates instructions, in real-time, to guide a user searching for a target object with a smartphone's camera. The solution is part of a more general active vision application for the visually impaired. An initial implementation of the system on an Android smartphone was experimentally evaluated with sighted participants to determine the performance of the control algorithm. The results show the effectiveness of our solution and its potential application to help the visually impaired find objects in unknown environments.}

@@

\noindent It is estimated that almost half a billion people worldwide live with mild to severe visual impairments or total blindness~\cite{bourne2017magnitude} and significant effort is being made to enable these people to lead more normal and independent lives. Modern improvements to mobile computing power and image processing techniques have provided researchers with new and powerful tools to solve this problem.
%The ActiVis project aims to contribute to this effort by delivering an indoor navigation system targeted towards enabling blind and visually impaired (VI) users to safely navigate towards their desired destination or
The work presented here is part of a project to assist the visually impaired (VI) navigate and find objects in unknown environments with the aid of a smartphone. The proposed system implements ideas from the field of active vision~\cite{Bajcsy2017}, but replaces the typical electro-mechanical actuators of a moving camera with the body (i.e.\ arm, hand) of the user holding the smartphone, as pictured in Figure~\ref{fig:system-use}. Schematically, the actuators used in a typical control loop, shown in Figure~\ref{fig:control-loop}, are replaced by a human. %This project expands upon concepts originally proposed in~\cite{bellotto2013} and~\cite{lock2017portable}.

\begin{figure}
  \centering
  \includegraphics[width=0.38\textwidth]{figures/system_use.png}
  \caption{The system in use during an experiment. }\label{fig:system-use}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{figures/control_loop.png}
  \caption{System control loop: $r$ is the reference object, $e$ the error signal, $u$ and $u^*$ the original and interpreted control signals and $y$ is the current object observation. $K$, $H$ and $P$ are the control, human and sensor blocks respectively. }\label{fig:control-loop}
\end{figure}

The goal of our active search system is to understand the user's surroundings and determine what the next best course of action is to reach the target object based on what is currently within view and what has been observed in the past. To this end, we implemented a smartphone's guidance system based on a Markov Decision Process (MDP)~\cite{bellman1957markovian} that generates, in real-time, an optimal set of instructions (policy) for the user to point to the target, depending on a previously-learned spatial distribution of known objects and on the current camera's view.  
%
This work includes three main contributions:

\begin{itemize}
  \item an MDP-based human controller that can guide a user in a visual search task;
  \item a data-based transition model for the MDP which includes spatial relations between known objects;
  \item a set of user experiments that prove the effectiveness of our active search implementation.
\end{itemize}

@@

Section~\ref{sec:previous-work} discusses other relevant work done in this field, followed by a general explanation of the active vision system for human guidance in Sections~\ref{sec:system-design}, and a detailed explanation of the human-control module in Section~\ref{sec:controller-design}. The experimental results are presented in Section~\ref{sec:experiments}, after which we conclude the paper and discuss future work in Section~\ref{sec:conclusion}.

@@

\noindent Assistive vision and guidance for the VI is a growing research area. In recent years, the increase in mobile processing power and computer vision improvements have led to research in the use of smartphone cameras to augment or enhance a user's vision and help them find objects or other points of interest. Earlier attempts at the problem involved placing special markers or barcodes around an environment, which the user then scans with a smartphone or similar mobile device~\cite{gude2013blind,iannizzotto2005badge3d,manduchi2012mobile}. This device then uses some feedback mode, e.g.~Braille or sound, to guide the user towards the target. %The major drawbacks to this approach is the significant upfront effort required to augment entire areas with markers to a point where the system will be useful to a VI user. Furthermore, these passive systems rely on the user scanning the marker in the first place; something that can be made more complicated with markers degrading over time, changing ambient lighting, different viewing angles and marker occlusion.

Another approach is to discard tags completely and rely on computer vision to perform the object detection, something that has become more practical with recent improvements to feature detectors and deep networks~\cite{huang2017speed,redmon2016you}. SIFT and SURF-based object detectors have also been used to detect known objects, when they are in the camera's view, and to guide the user to them using sonified instructions~\cite{schauerte2012assistive}. This type of systems is more flexible than the tag-based ones, but it has the same drawback of being {\em passive}, in the sense that it relies on having the object within the camera's view in the first place. Also, no clear performance metrics are reported in the previous paper. The VizWize system~\cite{bigham2010vizwiz} offloads the object recognition tasks to an Amazon Mechanical Turk worker who then provides feedback on where the object of interest is located relative to the user. The VizWiz has the advantage of being fairly robust and is able to classify a great deal of objects with little effort from the user and can provide natural, human-generated and curated directions. However, this approach does not enhance user independence, since the VI person is now beholden to an online worker instead of a relative, friend or bystander. Furthermore, a good internet connection is required on the device, possibly limiting its use in some poor-reception areas.

Many previous researchers have implemented active search and perception strategies in robots and image classifiers~\cite{Bajcsy2017} in an attempt to optimise their classification and planning tasks, for example by exploiting the structured nature of human environments and object placements. 
%
Two research teams have recently implemented an active object search strategy into their image classifiers~\cite{caicedo2015active,gonzalez2015active}. Their approaches use different methods but conceptually similar models to generate windows of interest for visual classification. The size and locations of the windows within the image are generated using the spatial relationship between objects, taken from the SUNCG and PASCAL datasets~\cite{song2016ssc,Everingham10}, and are iteratively changed based on the output from the respective models. The advantage of their approaches is that fewer windows are generated and submitted to the classifier, resulting in lower object classification times. The authors also report comparable accuracy compared to their counterparts.

Similar strategies have been incorporated on robotic platforms to improve autonomous object search, manipulation and localisation tasks. For example, some researchers have developed a planning algorithm for a robotic manipulator that performs an optimal object search in a cluttered environment~\cite{dogar2014object}. Another team implemented an MDP-based system generating an optimal object search strategy in a room over a belief state of object positions and configurations~\cite{aydemir2011search}. However, the authors trained their MDP using a custom object-placement and configuration scenario, so their results are sensitive to changes within this distribution.

In summary, much research has been conducted on recognition of and guidance towards target objects, including active vision solutions for image classifiers and robotic systems. However, to our knowledge, no previous work has been done on {\em active} object search and guidance for humans, which would especially benefit VI users. In this paper, we implement such an active vision system with a human in the loop
% ~\cite{bellotto2013,lock2017portable}
that guides the user towards an out-of-view target object. Our system exploits prior knowledge of the objects spatial distribution within an indoor environment, learned from a dataset of real-world images, and the history of past object observations made during the search.
%With this, we hope to make the searching process less random and move towards a more natural and efficient object searching process to the benefit of the VI\@.

@@

In comparison to the VizWiz system~\cite{bigham2010vizwiz} (mean 92s, standard deviation 37.7s), our results look very reasonable. However, these results may greatly change when we test the system with VI participants.
